---
title: "HW 2"
date: "September 26, 2019"
author: "Kelly Wang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Problem 1
## Load in Mr. Trash Wheel dataset

* Reads in the the dataset and specifically reads in the Mr. Trash Wheel Sheet
* Omits non-data entries (rows with notes/figures; columns containing notes)
* Uses function read_excel
* Uses Reasonable Variable names
* Omit Rows that do not include dumpset-specific data
* Round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)


```{r}
library(readxl)

##just want the "Mr. Trash Wheel" sheet and clean the data from there
waterwheel_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="Mr. Trash Wheel"
  ) %>%
  janitor::clean_names() %>%
  select(-x15) %>%
  drop_na() %>%
  mutate(
    sports_balls=as.integer(round(sports_balls, digits=0)) 
    ) %>%
  drop_na(dumpster)

#need to use reasonable variable names

view(waterwheel_data)
```
```{r}
## reading and cleaning Precipitation data for 2017. Adding year 2018 and omitting rows without preciptiation data  
Precip_2018_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="2018 Precipitation", range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2018) %>%
  filter(!is.na(total))

view(Precip_2018_data)
## reading and cleaning Precipitation data for 2018
Precip_2017_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="2017 Precipitation",
    range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2017) %>%
  filter(!is.na(total))

view(Precip_2017_data)

#stacking/binding to combine the two datasets 
month.name
twoyear_precip_data=
  bind_rows(Precip_2018_data, Precip_2017_data) %>%
  mutate(
    month = month.name[as.numeric(month)]
  )


view(twoyear_precip_data)
```

