---
title: "HW 2"
date: "September 26, 2019"
author: "Kelly Wang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Problem 1
## Load in Mr. Trash Wheel dataset

* Reads in the the dataset and specifically reads in the Mr. Trash Wheel Sheet
* Omits non-data entries (rows with notes/figures; columns containing notes)
* Uses function read_excel
* Uses Reasonable Variable names
* Omit Rows that do not include dumpset-specific data
* Round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)


```{r}
library(readxl)

##just want the "Mr. Trash Wheel" sheet and clean the data from there
waterwheel_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="Mr. Trash Wheel"
  ) %>%
  janitor::clean_names() %>%
  select(-x15) %>%
  drop_na() %>%
  mutate(
    sports_balls=as.integer(round(sports_balls, digits=0)) 
    ) %>%
  drop_na(dumpster)

view(waterwheel_data)
```
```{r}
## reading and cleaning Precipitation data for 2017. Adding year 2018 and omitting rows without preciptiation data  
Precip_2018_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="2018 Precipitation", range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2018) %>%
  filter(!is.na(total))

view(Precip_2018_data)
## reading and cleaning Precipitation data for 2018
Precip_2017_data=
  read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
    sheet="2017 Precipitation",
    range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2017) %>%
  filter(!is.na(total))

view(Precip_2017_data)

#stacking/binding to combine the two datasets 
month.name
twoyear_precip_data=
  bind_rows(Precip_2018_data, Precip_2017_data) %>%
  mutate(
    month = month.name[as.numeric(month)]
  )
view(twoyear_precip_data)
```
* median number of sports balls in dumpster 2017

In the Mr. Trash Wheel data, multiple information was taken. For example, the number of dumpsters that were filled per month, how much each dumpster weighed (in tons), what was in each dumpster, etc. It was found that between May 2014 and July 2018, a total of `r nrow(waterwheel_data)` dumpsters were collected. An example of an item that was found in each dumpster was sports balls. In the year 2017, the median number of sports balls in dumpster was `r median(pull(waterwheel_data,sports_balls, year==2017))`

It should also be noted that the amount of trash the device receives is highly dependent of rainfall. Dumpsters fill much faster when it rains than when it doesn't. Between 2017 and 2018, a total of `r nrow(twoyear_precip_data)` months were collected.  That being said, the total precipitation in 2017 was `r sum(pull(Precip_2017_data, total))` and the total precipitation in 2018 was `r sum(pull(Precip_2018_data, total))`. 

