---
title: "HW 2"
date: "October 2, 2019"
author: "Kelly Wang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
# Problem 1
## Load in Mr. Trash Wheel dataset

* Reads in the the dataset and specifically reads in the Mr. Trash Wheel Sheet
* Omits non-data entries (rows with notes/figures; columns containing notes)
* Uses function read_excel
* Uses Reasonable Variable names
* Omit Rows that do not include dumpset-specific data
* Round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)


```{r trashwheel}
library(readxl)

##just want the "Mr. Trash Wheel" sheet and clean the data from there
waterwheel_data=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet="Mr. Trash Wheel", range="A2:N408"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    sports_balls=as.integer(round(sports_balls, digits=0)) 
    ) %>%
  drop_na(dumpster)

view(waterwheel_data)
```
## Reading and Cleaning Precipitation data for 2017 and 2018
```{r precipitation}
## reading and cleaning Precipitation data for 2017. Adding year 2018 and omitting rows without preciptiation data  
Precip_2018_data=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet="2018 Precipitation", range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2018) %>%
  filter(!is.na(total))

view(Precip_2018_data)
## reading and cleaning Precipitation data for 2018
Precip_2017_data=
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet="2017 Precipitation",
    range="A2:B14"
  ) %>%
  janitor::clean_names() %>%
  mutate(year=2017) %>%
  filter(!is.na(total))

view(Precip_2017_data)

#stacking/binding to combine the two datasets 
twoyear_precip_data=
  bind_rows(Precip_2018_data, Precip_2017_data) %>%
  mutate(
    month = month.name[as.numeric(month)]
  )
view(twoyear_precip_data)
```
* median number of sports balls in dumpster 2017

In the Mr. Trash Wheel data, multiple information was taken. For example, the number of dumpsters that were filled per month, how much each dumpster weighed (in tons), what was in each dumpster, etc. It was found that between May 2014 and July 2018, a total of `r nrow(waterwheel_data)` dumpsters were collected. An example of an item that was found in each dumpster was sports balls. In the year 2017, the median number of sports balls in dumpster was `r filter(waterwheel_data, year==2017) %>% pull(sports_balls) %>% median()`

It should also be noted that the amount of trash the device receives is highly dependent of rainfall. Dumpsters fill much faster when it rains than when it doesn't. Between 2017 and 2018, a total of `r nrow(twoyear_precip_data)` months were collected.  That being said, the total precipitation in 2017 was `r sum(pull(Precip_2017_data, total))` and the total precipitation in 2018 was `r sum(pull(Precip_2018_data, total))`. 

# Problem 2
```{r fivethirtyeight}
# importing three datasets
#importing pols-month.csv and cleaning
pols_month_data =
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into =c("year", "month", "day"), sep = "-")%>%
  mutate(
    month = month.name[as.numeric(month)],
      year=as.integer(year)
  ) %>%
  pivot_longer(
    starts_with("prez"),
    values_to= "president"
  ) %>% 
  select(-day) %>% 
  arrange(year, month, desc(year)
  )
  
View(pols_month_data)
#importing snp.csv 
snp_data=
  read_csv(file="./data/fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  separate(col = date, into =c("month", "day", "year"), sep = "/")%>%
  select(year, month, day, close)%>%
  mutate(
    month = month.name[as.numeric(month)], 
      year=as.integer(year)
  ) %>% 
  arrange(year, month, desc(year)
  )
  
  
#importing unemployment.csv
unemployment_data=
  read_csv(file="./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names()%>%
  pivot_longer(
  jan:dec,
  names_to = "month",
  values_to="unemployment rate"
  ) %>% 
  mutate(
    month=recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr"= "April" , "may"= "May", "jun" = "June", "jul" = "July", "aug"="August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"), 
      year=as.integer(year)
) %>% 
  arrange(year, month, desc(year)
  )

view(unemployment_data)
view(snp_data)

#merging snp into pols, and then merging unemployment into the result
join_data = 
  left_join(pols_month_data, snp_data, by =c( "year","month"))%>%
  left_join(unemployment_data, by =c( "year","month"))

view(join_data)
```
The following datasets include rate of unemployment per specific months of a specific year, data on the number of natioanl politicians who were either democratic or republican within a given time, and information related to the stock market index (S&P). These three data sets were combined based off of the year and month, and were joined via left-join feature. The resulting data contains `r nrow(join_data)` rows and `r ncol(join_data)` columns. The resulting data set contains years between `r min(pull(join_data,year))` and `r max(pull(join_data,year))`. 

# Problem 3

* tidy data
* made sure categorical predictors are consistent
* makes sure all the case structures are the same (changed to all caps)
* remove duplicated items
```{r babynames}
library(dplyr)

baby_names_data=
  read_csv(file="./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "WHITE NON HISP" = "WHITE NON HISPANIC", "BLACK NON HISP" = "BLACK NON HISPANIC"),
    childs_first_name= tolower(childs_first_name)
  ) %>% 
  distinct() #removes duplicate rows of a dataframe
  

view(baby_names_data)
```
## Most Popular Female Baby Name Over Time
```{r}
olivia_df=
  filter(baby_names_data, childs_first_name =="olivia") %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

view(olivia_df)

knitr::kable(olivia_df)
```

## Most Popular Boy Baby Name Over Time 
```{r popular_baby_names}
finding_popular_boy_name_df=
  filter(baby_names_data, gender =="MALE", rank=="1")

#see that Jayden is the most common
pop_boy_name_df=
  filter(baby_names_data, childs_first_name=="jayden") %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
view(pop_boy_name_df)

knitr::kable(pop_boy_name_df)
```
## Creating a Scatter Plot Showing the Number of Children with a name (y axis) against the rank in popularity of that name (x axis)
```{r scatterplots}
scatter_df=
  filter(baby_names_data, gender== "MALE", ethnicity =="WHITE NON HISPANIC", year_of_birth == "2016")

ggplot(scatter_df, aes(x= childs_first_name, y=rank))
```
 
